\documentclass[a4paper,UKenglish]{lipics-v2016}
\usepackage{microtype}
\usepackage{amssymb,amsmath}
\usepackage{graphics,graphicx,color}
\usepackage{enumerate,paralist,hyperref}
\usepackage{caption,subcaption,xspace}
\usepackage{todonotes} 
\usepackage{times,setspace} 

%%%% Title of contribution
% ==================================================================
\title{Characterizations of Optimal 2- and 3-planar Graphs}
\titlerunning{Characterizations of Optimal 2- and 3-planar Graphs}
\author[1]{Michael~A.~Bekos}
\author[1]{Michael~Kaufmann}
\author[2]{Chrysanthi~N.~Raftopoulou}
% ==================================================================
\affil[1]{
Wilhelm-Schickhard-Institut f\"ur Informatik, Universit\"at T\"ubingen, Germany\\
\texttt{\{bekos,mk\}@informatik.uni-tuebingen.de}}
\affil[2]{
School of Applied Mathematical \& Physical Sciences, NTUA, Greece\\
\texttt{crisraft@mail.ntua.gr}}
% ==================================================================

% ==================================================================
\authorrunning{Michael~A.~Bekos, Michael~Kaufmann, Chrysanthi~N.~Raftopoulou}
% ==================================================================

\Copyright{Michael~A.~Bekos, Michael~Kaufmann, Chrysanthi~N.~Raftopoulou}
\subjclass{G.2.2 Graph Theory}% mandatory: Please choose ACM 1998 classifications from http://www.acm.org/about/class/ccs98-html . E.g., cite as "F.1.1 Models of Computation".
\keywords{$k$-planar graphs, characterization, graph drawing}%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Editor-only macros (do not touch as author)
\EventEditors{Boris Aronov, Matthew Katz}
\EventNoEds{1}
\EventLongTitle{33rd International Symposium on Computational Geometry}
\EventShortTitle{SoCG 2017}
\EventAcronym{SoCG}
\EventYear{2017}
\EventDate{July 4-7, 2017}
\EventLocation{Brisbane, Australia}
\EventLogo{}
\SeriesVolume{}
\ArticleNo{X} % Your paper number goes here! (not your EasyChair submission number)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newtheorem{property}{Property}

\newcommand{\pe}{potential edge\xspace}
\newcommand{\pes}{potential edges\xspace}
\newcommand{\pp}{potential empty cycle\xspace}
\newcommand{\pr}{potential region\xspace}
\newcommand{\pb}{potential boundary\xspace}

\newcommand{\parallelsum}{\mathbin{/\mkern-4mu/}}
\newcommand{\ppp}[2]{\angle [#1,#2]}
\newcommand{\ppc}[2]{\parallelsum [#1,#2]}

\begin{document}
\maketitle

% ============================================================================
\begin{abstract}
$k$-planarity is a concept that extends the class of planar graphs by allowing graphs that can be drawn with at most $k$ crossings on each edge. While for $1$-planar graphs, the \emph{optimal} graphs, i.e., those with the maximum number of edges, have been completely characterized, this was not the case for $k \geq 2$. For such cases, in particular for $k=2,3$ and $4$, upper bounds on the number of edges have been developed for the case of simple graphs by Pach and Toth, Pach et al.\ and Ackerman. Since then the bounds are known to be achieved by graphs with a very regular structure; the set of uncrossed edges form pentagonal or hexagonal faces respectively, which then are completely by $5$ more edges in the pentagons for the $2$-planar case, and by $8$ more edges in the hexagons for the $3$-planar case. Unfortunately, non-homotopic parallel edges have to be allowed. Previously, we have shown that the known bounds for the simple graphs hold also for the case of non-simple graphs with non-homotopic parallel edges and loops. In this paper, we prove that only those graphs achieve the bound of $5n-10$ for $2$-planar and the $5.5n -11$ bound for $3$-planar graph and provide a complete characterization of optimal non-simple $2$- and $3$-planar graphs.
\end{abstract}
% ============================================================================

\input{introduction.tex}
\input{preliminaries.tex}
\input{properties.tex}
\input{2planar.tex}
\input{3planar.tex}

% ============================================================================
\section{Conclusions}
\label{sec:conclusions}
% ============================================================================

In this paper, we developed techniques that allow us to characterize optimal $2$- and $3$-planar graphs. In particular, we had to deal with non-homotopic parallel edges, loops and also crossing adjacent edges, which required special attention and efforts. Natural extensions of our results would be to characterize optimal $4$-planar graphs, optimal fan-planar and quasi-planar graphs. But it would also be interesting to have a characterization of the simple versions of these graph classes where (non-homotopic) parallel edges and loops are forbidden. Another interesting line of research is the recognition of optimal $2$- and $3$-planar graphs. We note here the recent result by Brandenburg~\cite{DBLP:journals/corr/Brandenburg16a} who gives a linear time algorithm for the recognition of optimal $1$-planar graphs.

To Discuss:

\begin{itemize}
  \item Discussion on the desity of simple.
  \item Relation between optimal 2-planar and bar-1-visible.
  \item Relation between optimal 2-planar and 1-bend RAC.
\end{itemize}

% ============================================================================
\bibliographystyle{abbrv}
\bibliography{references}
% ============================================================================

\end{document}
